\documentclass[journal,twoside,web]{ieeecolor}
\usepackage{generic}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{algorithm,algorithmic}
\usepackage{hyperref}
\hypersetup{hidelinks=True}
\usepackage{textcomp}
\usepackage{subcaption}
%Added by Cheng
% \usepackage{multirow,array,enumitem,adjustbox}
\usepackage{booktabs,threeparttable,makecell}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\markboth{\hskip25pc IEEE TRANSACTIONS AND JOURNALS TEMPLATE}
{Author \MakeLowercase{\textit{et al.}}: Title}
\begin{document}
\title{Enhancing Alzheimer's Disease Diagnosis and Progression Prediction Through Multimodal Integration}
% \author{}
% \thanks{ is with the Department of (e-mail: ). }


\maketitle

\begin{abstract}
Alzheimer’s Disease (AD) classification methodologies increasingly use machine learning and deep learning. There is a gap in literature in using Vision Transformers (ViT) and hybrid vision transformer-convolutional neural net-vision transformers (ViT-CNN) in the multimodal prediction of AD using brain imaging. This study aims to: 1) classify AD status into 3 classes, 2) predict AD progression across 7 time-points, and 3) assess multimodal data integration's impact. Our approach combines CNNs and ViT for feature extraction from MRI, PET, EHR, and Genetic data, followed by feature fusion and classification. The hybrid ViT-CNN achieves 77\% validation and 70\% testing accuracy, outperforming individual models for AD classification. On the other hand, CNN (48-62 \%) outperforms the ViT (43-59 \%)  and the hybrid model (44-60 \%) for AD progression across seven timepoints on testing accuracy.  
Results indicate that PCA offers no benefit for AD progression classification, and SVM with 50 components performs best. Modality combination increases accuracy but decreases test sensitivity. Conclusion: The hybrid ViT-CNN model shows promise for AD classification not AD progression, extracting rich features from multimodal data. However, modality combination impacts sensitivity differently for CNN and ViT models.
\end{abstract}

\begin{IEEEkeywords}
Convolutional Neural Network, Vision Transformers, ADNI, Alzheimer's Disease, Support Vector Machine, Principal Component Analysis\
\end{IEEEkeywords}

\input{1-introduction}
\vspace{-3mm}

% \input{2-relatedwork}
% \vspace{-3mm}

\input{3-methodology}
\vspace{-3mm}

\input{4-results}
\vspace{-3mm}

\input{5-discussion}
\vspace{-3mm}

\input{6-conclusion}
\vspace{-3mm}

\section{References}
\bibliographystyle{ieeetr}
\bibliography{Bibliography}

\end{document}


